A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-03-18 paper/theor-experiment.tex
adding awa drawing"

%Currently we run the simulations on a low resolution space charge grid ($8
  %\times 8 \times 8$) as a proof of concept.
%In a next step a simple Python script would be applied to iteratively rerun
  %the optimizer with increasing space-charge resolution when the limits of the
  %design variables are narrow enough (iteratively diminishing search space). 
%\vspace{-7em}
\subsection{AWA Photoinjector Optimization}
Next we apply the optimization framework to the high charge beam line at the AWA, 
which consists of an rf photocathode gun, two solenoids, six linear accelerating cavities, 
followed by four quadrupoles and a stripline kicker, as shown in Fig. \ref{awa-pic}. 
Portions of this beam line served as the injector for Two Beam 
Acceleration (TBA) experiments in the past [X1], and will be used for 
similar experiments in the future [X2]. Main constraints on prior experimental 
results were the beam size and emittance as the beam passed through small aperture 
wakefield structures located after the quadrupoles. 
In an attempt to maximize charge transmission in upcoming experiments
we optimize several parameters in the gun and quadrupole leading into the 
TBA section of the beam line.
%\vspace{-1em}

@@

All simulations for this experiment were carried out on Bebop a
high performance computing (HPC)
cluster provided by the Laboratory Computing Resource Center (LCRC)
at Argonne National Laboratory (ANL). We used Intel Knights Landing 
(KNL) processors at 1.3 GHz with 128 GB of memory 
and 64 cores per node. There are 352 compute nodes available on 
Bebop, with a total of 22,528 cores. 
This allows for very large optimization jobs, like the AWA case.

OPAL was used as the forward solver, and a study on time step and 
number of particles was done to reduce the time to simulation while 
maintaining the physics of interest. All jobs were run and compared 
on 8 cores each, which allowed 8 jobs per node on the KNL.
The grid size was $16 \times 16 \times 32$, 
and parallelized in the x and y directions.
After comparing several options we choose ten thousand particles 
in combination with adjusting the time step near sensitive elements 
(i.e. quadurpoles, kicker). 
The resulting simulations are low fidelity, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig. \ref{tstep}.

@@

\begin{table}%[h!]
	\begin{center}
		\caption{Low and mid fidelity simulation parameters.}
		\label{fidelity}

		\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} C c D }
			\hline\noalign{\smallskip}
			& Number of particles & dT (seconds) & Time to simulation (minutes)\\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			low fidelity  			&  10,000   & $5 \times10^{-11}$  &  2.6 \\
			adjusted low fidelity  	&  10,000   & $2 \times 10^{-12}$, $1 \times 10^{-11}$ & 1.6 \\
			mid fidelity 			&  100,000  & $1 \times 10^{-12}$ &  18\\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

\subsubsection{AWA Optimization Results}

>>>>

%Currently we run the simulations on a low resolution space charge grid ($8
  %\times 8 \times 8$) as a proof of concept.
%In a next step a simple Python script would be applied to iteratively rerun
  %the optimizer with increasing space-charge resolution when the limits of the
  %design variables are narrow enough (iteratively diminishing search space). 
%\vspace{-7em}
\subsection{AWA Photoinjector Optimization} \label{awaproblem}
Next we apply the optimization framework to the high charge beam line at the AWA, 
which consists of an rf photocathode gun, two solenoids, six linear accelerating cavities, 
followed by four quadrupoles and a stripline kicker, as shown in Fig. \ref{awa-pic}. 
Portions of this beam line served as the injector for Two Beam 
Acceleration (TBA) experiments in the past [X1], and will be used for 
similar experiments in the future [X2]. Main constraints on prior experimental 
results were the beam size and emittance as the beam passed through small aperture 
wakefield structures located after the quadrupoles. 
In an attempt to maximize charge transmission in upcoming experiments
we optimize several parameters in the gun and quadrupole leading into the 
TBA section of the beam line.
%\vspace{-1em}

@@

All simulations for this experiment were carried out on Bebop a
high performance computing (HPC)
cluster provided by the Laboratory Computing Resource Center (LCRC)
at Argonne National Laboratory (ANL). We used Intel Knights Landing 
(KNL) processors at 1.3 GHz with 128 GB of memory 
and 64 cores per node. There are 352 compute nodes available on 
Bebop, with a total of 22,528 cores. All jobs were run and compared 
on 8 cores each, which allowed 8 jobs per node on the KNL.
This in combination with the amount of nodes available 
allows for very large optimization jobs, like the AWA case.

A study on time step and 
number of particles was done to reduce the time to simulation while 
maintaining the physics of interest. 
The grid size was $16 \times 16 \times 32$, 
and parallelized in the x and y directions.
After comparing several options we choose ten thousand particles 
in combination with adjusting the time step near sensitive elements 
(i.e. quadurpoles, kicker). 
The resulting simulations are low fidelity, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig. \ref{tstep}. For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
rms$_x$ and $\epsilon_x$ in the low fidelity results. A summary of 
the parameters used is listed in Table \ref{fidelity}.

@@

\begin{table}%[h!]
	\begin{center}
		\caption{Low and mid fidelity simulation parameters.}
		\label{fidelity}
		\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} C c D }
			\hline\noalign{\smallskip}
			& Number of particles & dT (seconds) & Time to simulation (minutes)\\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			low fidelity  			&  10,000   & $5 \times10^{-11}$  &  2.6 \\
			adjusted low fidelity  	&  10,000   & $2 \times 10^{-12}$, $1 \times 10^{-11}$ & 1.6 \\
			mid fidelity 			&  100,000  & $1 \times 10^{-12}$ &  18\\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

\subsubsection{AWA Optimization Results}
Given the parameters and variable boundaries in Section \ref{awaproblem}, 
four 50 generation optimization experiments were done with varying GA parameters. 
Similar to the time step scan, the goal was to determine which set of optimization
parameters strongly influence the results, and whether there was a time to solution
difference. (NOTE FOR EDIT: May need to take this out, or compare at 50 generations, none of these runs finished 100 gens)

\vspace{0.2cm}
{\footnotesize\begin{verbatim}
OPTIMIZE, INPUT="tmpl/optLinac_40nC.tmpl",
	OUTPUT="optLinac_40nC",
	OUTDIR="results",
	OBJECTIVES = {rmss1, rmsx1, emitx1, de1, rmss2, rmsx2, emitx2},
	DVARS = {dv0, dv1, dv2, dv3, dv4, dv5, dv6, dv7},
	CONSTRAINTS = {c1, c2},
	INITIALPOPULATION=656,
	MAXGENERATIONS=200,
	NUM_MASTERS=1,
	NUM_COWORKERS=8,
	SIMTMPDIR="tmp",
	TEMPLATEDIR="tmpl",
	FIELDMAPDIR="fieldmaps",
	NUM_IND_GEN=328,
	GENE_MUTATION_PROBABILITY=0.8,
	MUTATION_PROBABILITY=0.8,
	RECOMBINATION_PROBABILITY=0.2;
QUIT;
\end{verbatim}}
